---
title: "NOAA Stormwater Data Wrangle"
author: "Natalie Dornan"
date: "November 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

## Data Wrangling for Kahuwai Stormwater Data
## N Dornan

##Load Packages

library(tidyverse)
library(lubridate)
library(stringr)
library(tseries)

## Read in file

rain_dat = read.csv("NOAA_precipdata_KW_111719.csv")

#rain_dat$DATE <- ymd(rain_dat$DATE)

```

Tidy up the data to make it a little easier to work with
```{r}

wailupe_tidy <- rain_dat %>% 
  rename(station = STATION, station_name = STATION_NAME, elevation = ELEVATION, lat = LATITUDE, lon = LONGITUDE, date = DATE, time = TIME, qgag = QGAG, qgag_flag = Measurement.Flag, qgag_qual = Quality.Flag, qgag_units = Units, qpcp = QPCP, qpcp_flag = Measurement.Flag.1, qpcp_qual = Quality.Flag.1, qpcp_units = Units.1) %>%
  filter(station_name == "WAILUPE VALLEY SCHOOL 723.6 HI US") %>% #filter to Wailupe gauge only
  filter(qpcp != "-9999",
         qpcp != "999",
         qpcp != "999.99",
         qpcp_flag != "g",
         qgag != -9999.00) ## removes all flagged data
    
wailupe_pcp_sum <- wailupe_tidy %>%  
  group_by(date) %>%
  summarize(
    daily_gag = sum(qpcp)) ## gives total summed precip data per day. HT is given in tenths of inches.
```


Use lubridate to clean up the dates and times...

```{r}
##TS TIME

wailupe_pcp_sum$date <- ymd(wailupe_pcp_sum$date)

wailupe_ts <- wailupe_pcp_sum %>%
  ts(daily_gag, start=c(2008, 1), end=c(2014, 12), frequency=12)

plot(wailupe_ts)


######################## decompose ts

wailupe_dc <- decompose(wailupe_ts)

plot(wailupe_dc)

######################## standard ggplot

wailupe_plot <- ggplot(wailupe_pcp_sum, aes(date, daily_gag)) +
  geom_line()

wailupe_plot

#########Now, we filter by year! Need to reattach plyr to use str_detect

library(plyr)

#wailupe_refine = wailupe_dat %>%
#  filter(str_detect(date, paste(c(".2008", ".2009", ".2010"), collapse = #'|')))

wailupe_refine <- wailupe_pcp_sum %>%
  filter(date >"2008-01-01",
         date <"2010-12-31")

wailupe_ts_2 <- wailupe_refine %>%
  ts(wailupe_refine$qgag, start=c(2008, 1), end=c(2010, 12), frequency=12)

plot(wailupe_ts_2)

wailupe_plot_2 <- ggplot(wailupe_refine, aes(date, daily_gag)) +
  geom_line()

wailupe_plot_2

#write.csv(wailupe_refine, file = "wailupe_rainfall_r.txt")   
  
```

Awesome. The data is wrangled, now we need to pull out a good calibration sub dataset to feed into our model. 
To do this, Natalie will filter by year, and see the percentage of data present per year (#days in data/365). Then, she will choose a representative dataset from the resulting subset. 


```{r}



```

