---
title: "NOAA Stormwater Data Wrangle"
author: "Natalie Dornan"
date: "November 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}

## Data Wrangling for Kahuwai Stormwater Data
## N Dornan

##Load Packages

library(tidyverse)
library(lubridate)
library(stringr)
library(tseries)

## Read in file

rain_dat_77 = read.csv("NOAA_precipdata77_KW_02032020.csv")
rain_dat_08 = read.csv("NOAA_precipdata08_KW_111719.csv")
rain_paiko = read.csv("NOAA_paikodaily_KW_02062020.csv")

##Use lubridate to clean up the dates and times...

rain_dat_77$DATE <- ymd(rain_dat_77$DATE)
rain_dat_08$DATE <- ymd(rain_dat_08$DATE)
rain_paiko$DATE <- ymd(rain_paiko$DATE)

```

Tidy up the data to make it a little easier to work with
```{r}

wailupe_tidy_77 <- rain_dat_77 %>% 
  rename(station = STATION, station_name = STATION_NAME, elevation = ELEVATION, lat = LATITUDE, lon = LONGITUDE, date = DATE, time = TIME, qgag = QGAG, qgag_flag = Measurement.Flag, qgag_qual = Quality.Flag, qgag_units = Units, qpcp = QPCP, qpcp_flag = Measurement.Flag.1, qpcp_qual = Quality.Flag.1, qpcp_units = Units.1) %>%
  filter(station_name == "WAILUPE VALLEY SCHOOL 723.6 HI US") %>% #filter to Wailupe gauge only
  filter(qpcp != "-9999",
         qpcp != "999",
         qpcp != "999.99",
         qpcp_flag != "g",
         qpcp_flag != "{",
         qpcp_flag != "}",
         qpcp_flag != "[",
         qpcp_flag != "]",
         qgag != "-9999.00",
         qgag != "-9999",
         qgag_flag != "g",
         qgag_flag != "V",
         qgag_flag != "P",
         qgag_flag != "{",
         qgag_flag != "}",
         qgag_flag != "[",
         qgag_flag != "]") ## removes all flagged data
    
wailupe_daily_77 <- wailupe_tidy_77 %>%  
  group_by(date) %>%
  summarize(
    daily_pcp = sum(qpcp),
    daily_vol = sum(qgag)) ## gives total summed precip data per day. HT is given in hundreths of inches.


```



```{r}
##TS TIME

wailupe_daily_77$date <- ymd(wailupe_daily_77$date)

wailupe_ts <- wailupe_daily_77 %>%
  ts(daily_gag, start=c(1996, 1), end=c(2014, 12), frequency=12)

plot(wailupe_ts)


######################## standard ggplot

wailupe_plot_pcp <- ggplot(wailupe_daily_77, aes(date, daily_pcp)) +
  geom_line()

wailupe_plot_pcp

##search for 2 year, 24 hour storm event (4.78 inches plus/minus 4.12-5.57 inches 90% confidence interval!)

wailupe_investigate_1 <- wailupe_daily_77 %>%
  filter(daily_pcp > "4.12",
         daily_pcp < "5.57")

## Storms that fall in this range are on:
# 2004-01-02
# 2005-01-29
# 2010-12-19

## Look at scatter of qpcp vs qgag
scatterplot <- ggplot(wailupe_daily_77, aes(x= daily_pcp, y = daily_vol)) +
                        geom_point()

scatterplot

## filter dataset from 2004-2014
wailupe_clean_daily <- wailupe_daily_77 %>%
  filter(date > "2004-01-01",
         date < "2014-12-31")

## look at histogram plot to investigate storm frequency over given time period
hist <- ggplot(wailupe_clean_daily, aes(x= daily_pcp)) +
  geom_histogram()

hist

```

```{r}
#########Explore Paiko Dataset

paiko_precip <- rain_paiko %>%
  select(DATE,PRCP) %>%
  filter(PRCP > "0")

paiko_explore <- ggplot(paiko_precip, aes(x = DATE, y = PRCP)) +
  geom_point()

paiko_explore
  
```

Awesome. The data is wrangled, now we need to pull out a good calibration sub dataset to feed into our model. 
To do this, Natalie will filter by year, and see the percentage of data present per year (#days in data/365). Then, she will choose a representative dataset from the resulting subset. 


```{r} 

## This could be more elegant. Another way to do this is filtering by season.

######################
wailupe_05_investigate <- wailupe_tidy_77 %>%
  filter(date >"2005-1-1",
         date <"2005-12-31") %>%
  summarize(
    days_05 = length(date),
    annual_percent = (days_05/365)*100)

######################
wailupe_06_investigate <- wailupe_tidy_77 %>%
  filter(date >"2006-1-1",
         date <"2006-12-31") %>%
  summarize(
    days_06 = length(date),
    annual_percent = (days_06/365)*100)

######################3
wailupe_07_investigate <- wailupe_tidy_77 %>%
  filter(date >"2007-1-1",
         date <"2007-12-31") %>%
  summarize(
    days_07 = length(date),
    annual_percent = (days_07/365)*100)

######    
    
wailupe_08_investigate <- wailupe_tidy_77 %>%
  filter(date >"2008-1-1",
         date <"2008-12-31") %>%
  summarize(
    days_08 = length(date),
    annual_percent = (days_08/365)*100
                )

#################

wailupe_09_investigate <- wailupe_tidy_77 %>%
  filter(date >"2009-1-1",
         date <"2009-12-31") %>%
  summarize(
    days_09 = length(date),
    annual_percent = (days_09/365)*100
  )

#################
wailupe_10_investigate <- wailupe_tidy_77 %>%
  filter(date >"2010-1-1",
         date <"2010-12-31") %>%
  summarize(
    days_10 = length(date),
    annual_percent = (days_10/365)*100
  )

#################
wailupe_11_investigate <- wailupe_tidy_77 %>%
  filter(date >"2011-1-1",
         date <"2011-12-31") %>%
  summarize(
    days_11 = length(date),
    annual_percent = (days_11/365)*100
  )
#################
wailupe_12_investigate <- wailupe_tidy_77 %>%
  filter(date >"2012-1-1",
         date <"2012-12-31") %>%
  summarize(
    days_12 = length(date),
    annual_percent = (days_12/365)*100
  )

#################
wailupe_13_investigate <- wailupe_tidy_77 %>%
  filter(date >"2013-1-1",
         date <"2013-12-31") %>%
  summarize(
    days_13 = length(date),
    annual_percent = (days_13/365)*100
  )

```

filter years with targeted storm events

```{r}

wailupe_04 <- wailupe_daily_77 %>%
  filter(date >"2004-01-01",
         date <"2004-12-31")

wailupe_plot_04 <- ggplot(wailupe_04, aes(date, daily_pcp)) +
  geom_line() +
  xlab("Date") +
  ylab("Precipitation (inches)")

wailupe_plot_04

###############

wailupe_05 <- wailupe_daily_77 %>%
  filter(date >"2005-01-01",
         date <"2005-12-31")

wailupe_plot_05 <- ggplot(wailupe_05, aes(date, daily_pcp)) +
  geom_line() +
  xlab("Date") +
  ylab("Precipitation (inches)")

wailupe_plot_05


###############

wailupe_10 <- wailupe_daily_77 %>%
  filter(date >"2010-01-01",
         date <"2010-12-31")

wailupe_plot_10 <- ggplot(wailupe_10, aes(date, daily_pcp)) +
  geom_line() +
  xlab("Date") +
  ylab("Precipitation (inches)")

wailupe_plot_10



```


```{r}

##search for 2 year, 24 hour storm events (4.78 inches plus/minus 4.12-5.57 inches 90% confidence interval!)

wailupe_storm_04 <- wailupe_tidy_77 %>%
  filter(date > "2004-01-01",
         date < "2004-01-04")

storm_plot_04 <- ggplot(wailupe_storm_04, aes(time, qpcp)) +
  geom_point() +
  xlab("Time") +
  ylab("Precipitation (inches)") +
  theme_classic()

storm_plot_04

write.csv(wailupe_storm_04, file = "wailupe_storm_20040102_r.csv")  

######

wailupe_storm_05 <- wailupe_tidy_77 %>%
  filter(date > "2005-01-27",
         date < "2005-01-30")

storm_plot_05 <- ggplot(wailupe_storm_05, aes(time, qpcp)) +
  geom_point() +
  xlab("Time") +
  ylab("Precipitation (inches)") +
  theme_classic()

storm_plot_05

write.csv(wailupe_storm_05, file = "wailupe_storm_20050128_r.csv")  

##### condensed 2010 storm used for calibration

wailupe_storm_10_test <- wailupe_tidy_77 %>%
  filter(date == "2010-12-19")

storm_plot_10_test <- ggplot(wailupe_storm_10_test, aes(time, qpcp)) +
  geom_point() +
  xlab("Time") +
  ylab("Precipitation (inches)") +
  theme_classic()

storm_plot_10_test

write.csv(wailupe_storm_10_test, file = "wailupe_storm_10test_r.csv") 

#####

wailupe_storm_10 <- wailupe_tidy_77 %>%
  filter(date > "2010-12-18",
         date < "2010-12-20")

storm_plot_10 <- ggplot(wailupe_storm_10, aes(time, qpcp)) +
  geom_point() +
  xlab("Time") +
  ylab("Precipitation (inches)") +
  theme_classic()

storm_plot_10

write.csv(wailupe_storm_10, file = "wailupe_storm_20101219_r.csv") 

```

