---
title: "subcatchments_wailupe_final"
author: "Erica Johnson"
date: "5/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r , include=FALSE}

library(tidyverse)
library(dplyr)
library(data.table)
library(janitor)

points <- read_csv("subcatchment_points_all.csv") %>% 
  clean_names() %>% 
  rename(name = subcatchment_name,
         x = point_x,
         y = point_y
         ) %>%  
  select(name, x, y)

attributes <- read_csv("subcatchment_attributes.csv") %>% 
  clean_names() %>% 
  rename(name = objectid_1)
```

Collect Name, Area, Width, X and Y coordinates of each polygon and arrange them into the SWMM format. Rain gage and Outlets were assigned manually in SWMM

See example SWMM format below:

[SUBCATCHMENTS]								
;;Name  Rain Gage  Outlet  Area  %Imperv 	Width  %Slope  CurbLen  SnowPack      
        
```{r}
merge <- merge(points, attributes, by = "name" )

subc <- merge %>%
  mutate(
    area_acre = area_sqft*0.0000229568
    ) %>% 
  rename(
    area = area_acre,
    imperv = percent_imp
  ) %>% 
  distinct()


#add columns for "subcatchments" file. 

subc$rain_gage <- "R1" #placeholder until you can manually assign the correct rain gage to the correct subcatchment
subc$outlet <- "O1" #placeholder until you can manually assign the correct outlet to each subcatchment
subc$curblen <- 0

#arrange
subc_file <- subc %>% 
  select(
    name, 
    rain_gage, 
    outlet, 
    area, 
    imperv, 
    width, 
    slope, 
    curblen
    ) %>% 
  distinct(name, .keep_all = TRUE)

write.csv(subc_file ,"subcatchments.csv", row.names = FALSE)
```


[SUBAREAS]								
;;Subcatchment  N-Imperv  N-Perv  S-Imperv  S-Perv  PctZero  RouteTo  PctRouted 	
;;--------------	----------	----------	----------		
1	0.01	0.1	0.05	0.05	25	OUTLET    		
```{r}
suba <- subc_file %>% 
  select(name) %>% 
  rename (subcatchment = name) %>% 
  distinct()

suba$n_imperv <- 0.01 #manning's n for impervious surfaces - values from SWMM Manual

suba$n_perv <- 0.4 #manning's n for pervious (natural) surfaces. 0.4 for forested subcatchments. Later, identify urban subcatchments and input 0.15. - values from SWMM Manual  

suba$s_imperv <- 0.2 #impervious surface depth of depression storage (in) - values in SWMM Manual

suba$s_perv <- 0.3 #pervious (natural) surface depth of depression storage - values in SWMM manual

suba$pctzero <- 0
suba$RouteTo <- "OUTLET"

#n is for Manning's n, and s is for Depth of Depression Storage.

write.csv(suba ,"subareas.csv", row.names = FALSE)
```

[INFILTRATION]					
;;Subcatchment  	CurveNum  	          	DryTime   
;;--------------	----------	----------	----------
```{r}
infil <- subc %>% 
  select(name,
         curve_number
         ) %>% 
  rename (
    subcatchment = name
    ) %>% 
  distinct()

infil$Blank <- 0.5 #this is conductivity, however this may have been depreciated in the model
infil$DryTime <- 7 #units in days

write.csv(infil ,"infiltration.csv", row.names = FALSE)
```

[Polygons]		
;;Subcatchment  	X-Coord           	Y-Coord           
;;--------------	------------------	------------------
```{r}
polygons <- points %>% 
  select(
    name,
    x,
    y) %>% 
  rename(subcatchment = name
         ) %>% 
  distinct()

write.csv(polygons ,"polygons.csv", row.names = FALSE)
```
