---
title: "conduit_vertices"
author: "Erica Johnson"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read libraries and data file here
```{r, include=FALSE}
#Libraries
library(tidyverse)
library(dplyr)
library(data.table)
library(geosphere)
library(janitor)
library(kableExtra)


#Data with clean names
data <- read_csv("network_join2.csv") %>% clean_names()

```

Select columns with the data we need to use. Rename them for easy reference.
```{r}
conduits <- data %>% 
  select(
  objectid,
  point_x,
  point_y,
  elevation,
  subcatchment,
  roughness,
  type,
  diameter,
  width,
  height
  ) %>% 
  rename(
    name = objectid,
    x = point_x,
    y = point_y,
    subc = subcatchment,
    elevation = elevation,
    shape = type
    ) %>% 
  mutate (shape = str_replace_all(shape, "Reinforced Concrete Pipe", "CIRCULAR")) %>% 
  mutate (shape = str_replace_all(shape, "Box Culvert", "RECT_CLOSED")) %>% 
  mutate (shape = str_replace_all(shape, "Channel", "TRAPEZOIDAL")) %>% 
  mutate (shape = str_replace_all(shape, "Ditch", "RECT_OPEN")) %>% 
  mutate (shape = str_replace_all(shape, "Other", "RECT_OPEN")) %>% 
  distinct()


#Add the rest of the columns we need. SWMM does not take "NA" the default blank value for R, so we populate columns with 0 for now.
conduits$initflow <- 0
conduits$maxflow <- 0
conduits$length <- 0
conduits$inoffset <- 0
conduits$outoffset <- 0
```

"Length" provided by USGS is distance between xy points. This "length"" is not the actual length of the conduit because it does not take into consideration height (xyz), so we will use the difference between lower distance between xy points calcualte the actual length further down in the code.  

We will also re-calculate distance between xy points because different sources return  different values for some of the conduits and some conduits need to have this distance calculated anyway because it is blank. 

Step 1 - Create conduit names and assign conduits to node systematically, using stream subcatchment and arranged node order

```{r}
#Arrange by subc and elevation
arrange_subc <- conduits[with(conduits, order(name, -y, na.last=FALSE)),]

conduit_assign <- arrange_subc %>% 
   mutate(
    node1 = row_number()
    ) %>% 
  mutate(
      node2 = row_number()-1
      )
conduit_assign$con1 <- apply(
  conduit_assign,1,function(x) paste((x[16]), sep ="_",(x[1]))
  ) 

#remove space
conduit_assign$con1 <- gsub(
  '\\s+', '', conduit_assign$con1
  )

conduit_assign$con2 <- apply(
  conduit_assign,1,function(x) paste((x[17]),sep ="_",(x[1]))
  ) 

#remove space
conduit_assign$con2 <- gsub(
  '\\s+', '', conduit_assign$con2
  )
```



2. Arrange and reshape data 
Note: each conduit has start xy and end xy,as well as two different nodes so there are duplicate rows for each conduit
```{r}
con1 <- conduit_assign %>%  select(
  con1, 
  subc, 
  x, 
  y, 
  elevation, 
  node1, 
  roughness,
  shape,
  diameter,
  width,
  height)




con2 <- conduit_assign %>%  select(
  con2,
  subc, 
  x, 
  y, 
  elevation, 
  node1) %>% 
  rename(
    con1 = con2
    )

merge1<- merge(con1, con2, by = "con1") 
```

3. Length

We must now calculate the length of the conduits using the following steps:
a. Find distance "length" between xy coordinates of each conduit using geosphere.
b. Use difference in elevation to calculate height
c. Use pythag. theorem to calculate length

```{r}
#part a -  distance (adjust code based on number of pairs. This dataset has 18 based on the longest conduit)
dist <- merge1 %>%  rowwise() %>% mutate (dist_m = distm(c(x.x, y.x), c(x.y, y.y), fun=distHaversine)) %>% mutate(dist_ft = dist_m*3.28084)

#part b and c - height then length
lengths <- dist %>%  
  mutate (length = sqrt((dist_ft)^2 +(elevation.x-elevation.y)^2)
          )%>% 
  rename (
    name = con1,
    from_node = node1.x,
    to_node = node1.y,
    subc = subc.x
  )

```

4. file output for conduits
[CONDUITS]								
;;Name  From Node To Node  Length Roughness InOffset  OutOffset InitFlow  MaxFlow
;;---------- ---------- ---------- ---------- ---------- ---------- ---------- ---------- ----------

Use roughness value 0.01, for concrete pipes found in Appendix A-8 pg. 184 of EPA manual
```{r}
conduits <- lengths %>% 
  mutate(
    roughness = ifelse(is.na(roughness), 0.01, roughness)
    ) %>% 
  select(
    name,
    from_node, 
    to_node,
    length, 
    roughness 
    ) 
conduits$inoffset <- 0
conduits$outoffset <- 0
conduits$initflow <- 0
conduits$ maxflow <- 0


write.csv(conduits,"conduits.csv", row.names = FALSE)
```

5. file output for lengths
[XSECTIONS]							
;;Link  Shape	Geom1	Geom2	Geom3 Geom4 Barrels Culvert   
;;--------------	------------	----------------	----------

a - concrete pipe dimensions
```{r}
pipes<- lengths %>% 
  filter(
    shape == "CIRCULAR"
    ) %>% 
  rename(
    geom1 = diameter,
    link = name
    ) %>%
  mutate(
    geom1 = ifelse(is.na(geom1), "23", geom1)
    ) %>% 
  mutate(
    geom1 = ifelse((geom1=="Other"), "23", (geom1))
    )%>% 
  select(
    link, 
    shape,
    subc,
    length,
    geom1
  )
pipes$geom2 <- 0
pipes$geom3 <- 0
pipes$geom4 <- 0
pipes$barrels <- NA
pipes$culvert <- 0
```

b - box culverts, ditches, and "other" conduit dimensions
```{r}
ditch_box <- lengths %>% 
  filter(
    shape == "RECT_CLOSED" | shape == "RECT_OPEN"
    ) %>% 
  rename (
    geom1 = width, 
    geom2 = height,
    link = name
    ) %>% 
  mutate(
    geom1 = ifelse(is.na(geom1), 5, geom1)
    ) %>% 
  mutate(geom2 = ifelse(is.na(geom2), 5, geom2)
         ) %>% 
  select( 
    link, 
    shape, 
    subc,
    length,
    geom1,
    geom2
    
  )
ditch_box$geom3 <- 0
ditch_box$geom4 <- 0
ditch_box$barrels <- NA
ditch_box$culvert <- 0
```

c - channel (channelized stream) dimenssions
```{r}
channel <- lengths %>% 
  filter(
    shape == "TRAPEZOIDAL"
    ) %>% 
  rename(
    geom1 = width, 
    geom2 = height,
    link = name
    ) %>% 
  mutate(
    geom1 = ifelse(is.na(geom1), 30, geom1)
    ) %>% 
  mutate(
    geom2 = ifelse(is.na(geom2), 10, geom2)
    ) %>% 
  select(
     link, 
    shape,
    subc,
    length,
    geom1,
    geom2
    )
#Geom3 and Geom4 are side lopes, which literature indicate vary from 1/1 to 1/2.  Ive seen side slopes perpendicular to the ground, especially where homes are built.
channel$geom3 <- 1
channel$geom4 <- 1
channel$barrels <- 1
channel$culvert <- 0
```
e- bind all tables

```{r}
xsections_df <- rbind(pipes, ditch_box, channel) %>% select (link, shape, geom1, geom2, geom3, geom4, barrels, culvert, subc, length) %>%  distinct(link, shape, geom1, geom2, .keep_all = TRUE)

xsections <- rbind(pipes, ditch_box, channel) %>% select (link, shape, geom1, geom2, geom3, geom4, barrels, culvert) %>%  distinct(link, shape, geom1, geom2, .keep_all = TRUE)

write.csv(xsections,"xsections.csv", row.names = FALSE)
```

8. Junctions and coordinates/vertices

[JUNCTIONS]					
;;Name  Elevation MaxDepth  InitDepth SurDepth  Aponded   
;;--------------	----------	----------	----------	
```{r}

junctions <- conduit_assign %>% select(node1, elevation) %>% rename(name = node1)
junctions$maxdepth <- 0
junctions$initdepth <- 0
junctions$surdepth <- 0
junctions$aponded <- 0

write.csv(junctions,"junctions.csv", row.names = FALSE)
```

[COORDINATES]		
;;Node          	X-Coord           	Y-Coord           
;;--------------	------------------	------------------

SWMM takes x and y that are in decimal degrees  (lat and long).
```{r}
coordinates <- conduit_assign %>% select(node1, x, y) %>% rename(x_coord = x, y_coord =y)

write.csv(coordinates,"coordinates.csv", row.names = FALSE)
```
We will have to decide which points are "vertices" which are points of change in conduit direction and which are "junctions" which are points where two or more conduits join.


Step 7 - make a table by subcatchment for total lengths and average diameter/width/height
```{r}
pipe = xsections_df %>% 
  filter (shape =="CIRCULAR") %>% 
  group_by(subc) %>% 
  summarize(
    sum_length_pipe = sum(length),
    mean_diam_pipe= mean(as.numeric(geom1))
    )

 ditch = xsections_df %>% 
  filter (shape =="RECT_OPEN") %>% 
  group_by(subc) %>% 
  summarize(
    sum_length_ditch = sum(length),
    mean_width_ditch= mean(as.numeric(geom1)),
    mean_height_ditch= mean(as.numeric(geom2))
    )
 
 box = xsections_df %>% 
  filter (shape =="RECT_CLOSED") %>% 
  group_by(subc) %>% 
  summarize(
    sum_length_box = sum(length),
    mean_width_box= mean(as.numeric(geom1)),
    mean_height_box= mean(as.numeric(geom2))
    )
 
 channel = xsections_df %>% 
  filter (shape =="TRAPEZOIDAL") %>% 
  group_by(subc) %>% 
  summarize(
    sum_length_channel = sum(length),
    mean_width_channel= mean(as.numeric(geom1)),
    mean_height_channel= mean(as.numeric(geom2)),
    mean_slope1_channel = mean(as.numeric(geom3)),
    mean_slope2_channel = mean(as.numeric(geom4))
    )
```

```{r}

write.table(pipe, file="total_lengths.csv", sep=",", eol="\n", append=TRUE, row.names=FALSE, col.names=TRUE)
write.table(ditch, file="total_lengths.csv", sep=",", eol="\n", append=TRUE, row.names=FALSE, col.names=TRUE)
write.table(box, file="total_lengths.csv", sep=",", eol="\n", append=TRUE, row.names=FALSE, col.names=TRUE)
write.table(channel, file="total_lengths.csv", sep=",", eol="\n", append=TRUE, row.names=FALSE, col.names=TRUE)


```
