---
title: "Error115"
author: "Erica Johnson"
date: "11/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file will address the adverse conduit slopes error from SWMM

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
library(geosphere)


```

```{r, include=FALSE}
conduits_data <- read_csv("Conduit_XY_Elev_SpatialJoin.csv") %>% rename(Name = OBJECTID)
error_115 <- read_csv("trouble_conduits_erica.csv") %>% rename(Name = Conduit)
SWMM_arrangement <- read_csv("conduits_in_SWMM.csv") # had to remove spaces between From Node and To Node columns. We should use snakecase for this later....
```

Match the original conduit data to error conduits. The match function removes all non-matches. I will use this match to take a looke at the nodes that match each conduit, and their elevation
```{r}
merge_1 <-merge(conduits_data, error_115, by = "Name", all = TRUE) %>%  select(Name, Error, OBJECTID_12, RASTERVALU_1, POINT_X, POINT_Y ) %>% rename (Nodes = OBJECTID_12, Elevation = RASTERVALU_1)
#adverse slope error is due to negative slope because flow is being directed to a higher elevation.This is because we do not know invert elevation and used elevation as a proxy. When checking conduit flow in the model, flow was directed towards central conduits and towards the outfall (ocean)

merge_2 <- merge(merge_1, SWMM_arrangement, by = "Name", all = TRUE)
#now we match the error conduits to the ones in the SWMM input file. This is so we know the order of the to and from nodes. After a quick look, it does seem to be the case that every "to" node is at a higher elevation than the "from" nodes

#arrange be Name and Elevation. This was the same arrangement used for conduits and should replicate the arrangement of the "FromNode" and "ToNode"
arrange_nodes <- merge_2[with(merge_2, order(Name, Elevation, na.last=FALSE)),]

match_elev <- dcast(setDT(arrange_nodes), Name~rowid(Name, prefix="Elevation"), value.var="Elevation")
#Elevation1 is the elevation of the FromNode, and Elevation 2 is the elevation of the ToNode

#merge data frames
merge_3 <- merge(merge_2, match_elev, by = "Name")


```

```{r}
#arrange again for good measure
arrange_nodes2 <- merge_3[with(merge_3, order(Name, Elevation, na.last=FALSE)),]

#add a column that repeats 1-2. anything stormwater structure with a 2 is a "ToNode"
arrange_nodes2$FromTo <-rep(1:2, length.out = 1709 )

select_nodes <-arrange_nodes2 %>% select (Name, Nodes, Error, Elevation, Elevation1, Elevation2, FromTo) %>% distinct() %>% mutate(criteria = FromTo*Error)

```

The minimm elevation drop across the link must be 0.001 /3.048 meters, or converted to 0.001/10.00000032 ft. Internally Elevation1 – Elevation2 should be greater than 0.001 /3.048 meters

```{r}
#create function
new_el<- function(x){x-(0.001/ (3.048*3.28084))}
old_el<- function(x){x}


#use mutate, case_when, and functions to replace elevation values for conduits that returned an error and keep elevation values for conduits that did not.
error_fix <- select_nodes %>% mutate(
 new_elev= case_when(
   criteria %in% c("230") ~ new_el(Elevation1-(0.001/(3.048*3.28084))),
   criteria %in% c(NA) ~ old_el(Elevation),
   criteria %in% c("115") ~ old_el(Elevation)
                     )) 

```

Some nodes have mutiple conduits feeding into them, so there may be multiple values assigned to a single node! We want to use the lowest elevation values to ensure both conduits can direct their flow to the right node. 
```{r}
#usually when duplicates are removes, the first in the list is kept. We want to arrange by nodes, then from lower to high elevations.
arrange <- error_fix[with(error_fix, order(Nodes, new_elev, na.last=FALSE)),]

#select the columns we want
elev_fix <- arrange %>%  select(Name, Nodes, Elevation1, new_elev) 

#remove duplicates by the "Nodes" column only
elev_fix1 <-elev_fix[!duplicated(elev_fix[,c("Nodes")]),] 
```


```{r}

Error115_nodes <- elev_fix1 %>% select(Nodes, new_elev)
write.csv(Error115_nodes ,"Error115_nodes.csv", row.names = FALSE)
```



New elevation will likely cause a conduit length error, so we will need to fix this too.
```{r}
#make a table with both x and y coordinates in one row

match_x <- dcast(setDT(merge_3), Name~rowid(Name, prefix="POINT_X"), value.var="POINT_X")

match_y <- dcast(setDT(merge_3), Name~rowid(Name, prefix="POINT_Y"), value.var="POINT_Y")


#The match function will only have the two variables. We will need to merge conduit and "match"" data tables together using the merge function.

merge_XY <-merge(match_x, match_y, by = "Name", all = TRUE)

merge_elev1 <-merge(merge_XY, elev_fix1, by = "Name", all = TRUE)


width <- merge_elev1 %>%  rowwise() %>% mutate (width = distm(c(POINT_X1, POINT_Y1), c(POINT_X2, POINT_Y2), fun=distHaversine)) %>% mutate(width_ft = width*3.28084)

length <- width %>%  mutate (length = sqrt(width_ft^2 +(Elevation1 - new_elev)^2)) %>% select ( Name, length) %>% filter(length !="NA") %>% distinct()

```

We are supposed to have 817 unique conduits

```{r}
write.csv(length ,"Error115_lengths.csv", row.names = FALSE)
```

