---
title: "stream_network"
author: "Erica Johnson"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#Libraries
library(tidyverse)
library(dplyr)
library(data.table)
library(geosphere)
library(janitor)
library(kableExtra)


#Data with clean names
#Data are conduits matched to structures. This is because we need conduit xsections from structures)
data1 <- read_csv("stream_elev.csv") %>% clean_names()
data2 <- read_csv("stream_poly.csv") %>% clean_names()
```

```{r}

stream1 <- data1 %>% 
  select(
  objectid_1,
  objectid, 
  point_x, 
  point_y, 
  rastervalu
  ) %>% 
  rename(
    nodes = objectid_1,
    name = objectid,
    elevation = rastervalu, 
    x = point_x, 
    y = point_y
    ) 

stream2 <- data2 %>% 
  select(
  objectid_1,
  rastervalu
  ) %>% 
  rename(
    nodes = objectid_1,
    subc = rastervalu
    ) 

streams <- merge(
  stream1, 
  stream2, 
  by = "nodes"
  )
```


1. Nodes are arranged in the order they need to be connected:
First sort them by decending y values then ascending X values
```{r}
arrange <-streams[with(streams, order(subc, -y, na.last=FALSE)),]
```

2. Create conduit names and assign conduits to node systematically, using stream subcatchment and arranged node order
```{r}

conduit_assign <- arrange %>% 
  mutate(
    count_nodes = row_number()
  ) %>% 
  group_by(
    subc
    ) %>%
  mutate(
    con_assign1 = row_number()
    ) %>% 
  mutate(
      con_assign2 = row_number()-1
      )

conduit_assign$nodes <- apply(
  conduit_assign,1,function(x) paste((x[7]), sep ="c",(x[2]))
  )
conduit_assign$con1 <- apply(
  conduit_assign,1,function(x) paste((x[2]), sep ="c",(x[8]))
  )

#remove space
conduit_assign$con1 <- gsub(
  '\\s+', '', conduit_assign$con1
  )

conduit_assign$con2 <- apply(
  conduit_assign,1,function(x) paste((x[2]),sep ="c",(x[9]))
  ) 

#remove space
conduit_assign$con2 <- gsub(
  '\\s+', '', conduit_assign$con2
  )

```

3. Arrange and reshape data 
Note: each conduit has start xy and end xy,as well as two different nodes so there are duplicate rows for each conduit
```{r}

con1 <- conduit_assign %>%  select(con1, subc, x, y, elevation, nodes)

con2 <- conduit_assign %>%  select(con2,subc, x, y, elevation, nodes) %>% rename(con1 = con2)

merge1 <- merge(con1, con2, by = "con1") 

```

4. We must now calculate the length of the conduits using the following steps:

a. Find distance "length" between xy coordinates of each conduit.
b. Use difference in elevation to calculate height
c. Use pythag. theorem to calculate length

a. Distance
```{r}

dist <- merge1 %>%  rowwise() %>% mutate (dist_m = distm(c(x.x, y.x), c(x.y, y.y), fun=distHaversine)) %>% mutate(dist_ft = dist_m*3.28084)

```

b-c : Height and Length
```{r}
length <- dist %>%  
  mutate(
    length = sqrt(dist_ft^2 +(elevation.x-elevation.y)^2)
         ) %>% 
  rename (
    name = con1,
    from_node = nodes.x,
    to_node = nodes.y,
    subc = subc.x
  )
```

```{r}
#add columns for Geom1, Geom2, Geom3, Geom4,Barrels, and Culvert
length$shape <- "Trapezoidal"
length$geom1 <- 30
length$geom2 <- 10
length$geom3 <- 1
length$geom4 <- 1
length$barrels <- "NA"
length$culvert <- 0
#added the model default values to columns. Not sure what they mean, but no error message from model now


```
4. file output for conduits

[CONDUITS]								
;;Name  From Node To Node  Length Roughness InOffset  OutOffset InitFlow  MaxFlow
;;---------- ---------- ---------- ---------- ---------- ---------- ---------- ---------- ----------
We used roughness value 0.07, the higher end of the range for natural channels found in Appendix A-8 pg. 185 of EPA manual

Using the subcatchments delineations, we filtered just the natural streams. The channelized stream in the urban area is represented by "channel"
```{r}
conduit_stream <- length %>% 
  filter(
    subc <= 3|subc == 34|subc == 38 |subc == 24 | subc == 7
  ) %>% 
  select(
    name,
    from_node, 
    to_node,
    length
    ) 
natural_stream$roughness <- 0.07
natural_stream$inoffset <- 0
natural_stream$outoffset <- 0
natural_stream$initflow <- 0
natural_stream$maxflow <- 0


write.csv(conduit_stream,"conduit_stream.csv", row.names = FALSE)
```

5. file output for xsection

[XSECTIONS]							
;;Link  Shape	Geom1	Geom2	Geom3 Geom4 Barrels Culvert   
;;--------------	------------	----------------	----------

```{r}
xsection_stream <- length %>% 
  filter(
    subc <= 3|subc == 34|subc == 38 |subc == 24 | subc == 7
  ) %>% 
  rename(
    link = name
    ) %>% 
  select(
     link, 
    shape
    )
xsection_stream$geom1 <- 30
xsection_stream$geom2 <- 10
xsection_stream$geom3 <- 1
xsection_stream$geom4 <- 1
xsection_stream$barrels <- 1
xsection_stream$culvert <- 0

write.csv(xsection_stream,"xsection_stream.csv", row.names = FALSE)
```

6. Junctions and Coordinates/Vertices

[JUNCTIONS]					
;;Name  Elevation MaxDepth  InitDepth SurDepth  Aponded   
;;--------------	----------	----------	----------	
```{r}

junctions_stream <- conduit_assign %>% 
  select(
    nodes, 
    elevation
    ) %>% rename(
      name = nodes
      ) %>% 
  filter(
    subc <= 3|subc == 34|subc == 38 |subc == 24 | subc == 7
  ) 
junctions_stream$maxdepth <- 0
junctions_stream$initdepth <- 0
junctions_stream$surdepth <- 0
junctions_stream$aponded <- 0

write.csv(junctions_stream,"junctions_stream.csv", row.names = FALSE)
```
[COORDINATES]		
;;Node          	X-Coord           	Y-Coord           
;;--------------	------------------	------------------

SWMM takes x and y that are in decimal degrees  (lat and long).
```{r}
coordinates_stream <- conduit_assign %>% 
  select(
    nodes, 
    x, 
    y
         ) %>% 
  rename(
    x_coord = x, 
    y_coord =y
    ) %>% 
  filter(
    subc <= 3|subc == 34|subc == 38 |subc == 24 | subc == 7
  ) 

write.csv(coordinates_stream,"coordinates_stream.csv", row.names = FALSE)
```
We will have to decide which points are "vertices" which are points of change in conduit direction and which are "junctions" which are points where two or more conduits join.


Step 7 - make a table by subcatchment for total lengths and average diameter/width/height

```{r}

#just select the upper watershed subcatchment. In the urban area, the stream is represented as "Channel".
stream_lengths = length %>% filter(subc <= 3|subc == 34|subc == 38 |subc == 24 | subc == 7) %>% 
  group_by(subc) %>% 
  summarize(
    sum_length_stream = sum(length),
    mean_width_stream= mean(as.numeric(geom1)),
    mean_height_stream= mean(as.numeric(geom2)),
    mean_slope1_stream = mean(as.numeric(geom3)),
    mean_slope2_stream = mean(as.numeric(geom4))
    )


```
```{r}
write.table(stream_lengths, file="total_lengths.csv", sep=",", eol="\n", append=TRUE, row.names=FALSE, col.names=TRUE)
```

