---
title: "streams"
author: "Erica Johnson"
date: "12/8/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
library(geosphere)
library(janitor)
```

For streams we must characterize them in SWMM as storage, conduits, and junctions. For now, we have XY coordinates and elevation for points along the stream. Elevation is how SWMM knows where the water should flow in a gravity fed system. The Elevation column in SWMM is actually "invert elevation".

See SWMM example below:

[JUNCTIONS]					
;;Name  Elevation MaxDepth  InitDepth SurDepth  Aponded   
;;--------------	----------	----------	----------	
22503	0	0	0	0	8.5


Reead in data
```{r, include = FALSE}

stream_data <- read_csv("stream_XY_elev.csv") 
 
```

Decisions regarding formatting:

There are duplicate Object IDs corresponding to a segment of the stream. We will use FID instead which are unique to each coordinate and should not pose any issues with the names of other points already in SWMM.

Raster Value is elevation in ft.


Select columns and names to use. We are working with conduits and structures in this data set, so conduit names = Names and node (strucutre) names = Nodes
```{r}

streams <- stream_data %>%  select( FID, RASTERVALU, POINT_X, POINT_Y) %>% rename(Nodes = FID, Elevation = RASTERVALU) 

```

Add columns
```{r}
streams$MaxDepth <- 0
streams$InitDepth <- 0
streams$SurDepth <- 0
streams$Aponded <- 0
```

Select columns
```{r}
stream_junctions <- streams %>% select( Nodes, Elevation,  MaxDepth, InitDepth, SurDepth, Aponded) %>%  distinct()

```


```{r}
write.csv(stream_junctions,"Wailupe_stream_junctions_SWMM.csv", row.names = FALSE)
```

Next we need to make a file with the xy coordinates of the structures so SWMM knows where to put them.

See SWMM example below:
[COORDINATES]		
;;Node          	X-Coord           	Y-Coord           
;;--------------	------------------	------------------
22503	1720700.042	41573.305
```{r}

#SWMM columns we need are: Node	X-Coord	Y-Coord
stream_coordinates <- stream_data %>% select (FID, POINT_X, POINT_Y) %>% rename (Name = FID ,X_Coord =  POINT_X , Y_Coord = POINT_Y) %>% distinct()

```

```{r}
write.csv(stream_coordinates,"Wailupe_stream_coordinates_SWMM.csv", row.names = FALSE)
```

Now we need to make stream conduits to connect the junctions.

1. Nodes are arranged in the order they need to be connected:
First sort them by decending y values then ascending X values
```{r}
arrange <-stream_data[with(stream_data, order(PEREN_ID, -POINT_Y, na.last=FALSE)),]
```


2. Create conduit names and assign conduits to node systematically, using stream perrenial ID and previously arranged node order
```{r}

conduit_assign <- arrange %>% group_by(PEREN_ID) %>%
  mutate(con_assign1 = row_number()) %>% mutate(con_assign2 = row_number()-1)


conduit_assign$con1 <- apply(conduit_assign,1,function(x) paste((x[4]), sep ="_",(x[14])))
#remove space
conduit_assign$con1 <- gsub('\\s+', '', conduit_assign$con1)

conduit_assign$con2 <- apply(conduit_assign,1,function(x) paste((x[4]),sep ="_",(x[15]))) 
#remove space
conduit_assign$con2 <- gsub('\\s+', '', conduit_assign$con2)
```




```{r}

df_con1 <- conduit_assign %>%  select(con1, FID, POINT_X, POINT_Y, RASTERVALU, PEREN_ID)

df_con2 <- conduit_assign %>%  select(con2, FID, POINT_X, POINT_Y, RASTERVALU, PEREN_ID) %>% rename(con1 = con2)

merge1 <- merge(df_con1, df_con2, by = "con1") 

```



We must now calculate the true length of the conduits using the following steps:

1. Find distance "length" between xy coordinates of each conduit.

2. Use difference in elevation to calculate height

3. Use pythag. theorem to calculate length

STEP 1. Distance
```{r}

XY_dist <- merge1 %>%  rowwise() %>% mutate (width = distm(c(POINT_X.x, POINT_Y.x), c(POINT_X.y, POINT_Y.y), fun=distHaversine)) %>% mutate(width_ft = width*3.28084)

```

Step 2-3 : use difference in elevation to calculate height, then length


```{r}

length <- XY_dist %>%  mutate (length = sqrt(width_ft^2 +(RASTERVALU.x-RASTERVALU.y)^2))


```


Select the columns we want for SWMM 

I used roughness value 0.07, the higher end of the range for natural channels found in Appendix A-8 pg. 185 of EPA manual

```{r}

# data frame for both file outputs:
streams <- length %>%  select(con1, FID.x, FID.y, length) %>% rename (Name = con1, FromNode = FID.x , ToNode = FID.y)
streams$Roughness <- 0.07
streams$InOffset <- 0
streams$OutOffset <- 0
streams$InitFlow <- 0
streams$MaxFlow <- 0

```

Condiut file output:
```{r}
write.csv(streams,"Wailupe_stream_conduits_SWMM.csv", row.names = FALSE)
```
Note: After inputting file data to SWMM, I deleted two conduits ( 399_34 and 422_184). Conduits were the result of nodes with the exact same coordinates (264 & 335, 336 & 542). 

	 
	 	
	 
	 
	 

The next file we will prepare are the cross sections of the conduits. See SWMM example below:
[XSECTIONS]							
;;Link  Shape	Geom1	Geom2	Geom3 Geom4 Barrels Culvert   
;;--------------	------------	----------------	----------
13108	Circular    	0	0	0	0	0	0

We will use Trapezoidal (open top) conduit type for the stream ass opossed to irregular natural channel because it has been channelized and is managed with a buldozer, and has mason ston retaining walls on each side. 

We can use the irregular channel for upper watershed conduits, but that will need to be identified by perrenial ID
```{r}

xsections <- streams %>% select(Name)

#add columns for Geom1, Geom2, Geom3, Geom4,Barrels, and Culvert
xsections$Type <- "Trapezoidal"
xsections$Geom1 <- 1
xsections$Geom2 <- 1
xsections$Geom3 <- 0
xsections$Geom4 <- 0
xsections$Barrels <- 1
xsections$Culvert <- 0

#added the model default values to columns. Not sure what they mean, but no error message from model now
                          
```

Xsections file output
```{r}

write.csv(xsections,"Wailupe_stream_xsections_SWMM.csv", row.names = FALSE)
   
```
