---
title: "sub catchments"
author: "Erica Johnson"
date: "11/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
```{r , include=FALSE}

library(tidyverse)
library(dplyr)
library(data.table)
library(rgdal) #this is for converting UTM to lat long. this is also not working!
```

I will collect Name, Area, Width, X and Y coordinates and arrange them into the SWMM format. Much of the data regarding these polygons will resort to the project default values because they are not given. 

See example SWMM format below:

[SUBCATCHMENTS]								
;;Name  Rain Gage Outlet Area %Imperv 	Width   	%Slope  	CurbLen 	SnowPack        
;;--------------	----------------	----------------	
1	*               	*               	5	25	500	0.5	0	                
								
```{r}
subcatch_data<- read_csv("subcatch_XY.csv")
```

```{r}

subcatch <- subcatch_data %>%  select(OBJECTID, Area_Acre, Shape_Leng, Percent_Ar, AveSlope,  POINT_X, POINT_Y) %>% 
  rename(Name = OBJECTID, Width = Shape_Leng, Imperv = Percent_Ar, Slope = AveSlope, X_Coord = POINT_X, Y_Coord = POINT_Y ) %>% mutate(Area = Area_Acre*43560)

#Convert acres to sq ft. I believe Shape_Leng is already measured in feet, so no need to convert. Ave slope values already looks like %. May need to change this if I am wrong

#add columns for "subcatchments" file, add model default values
subcatch$Rain_gage <- "R1"
subcatch$Outlet <- "O1"
subcatch$CurbLen <- 0


#arrange
subc <- subcatch %>% select(Name, Rain_gage, Outlet, Area, Imperv, Width, Slope, CurbLen) %>% distinct()
```


[SUBAREAS]								
;;Subcatchment  	N-Imperv  	N-Perv    	S-Imperv  	S-Perv  	PctZero   	RouteTo   	PctRouted 	
;;--------------	----------	----------	----------		
1	0.01	0.1	0.05	0.05	25	OUTLET    		
```{r}
suba <- subcatch %>% select(Name) %>% rename (Subcatchment = Name)%>% distinct()
suba$N_Imperv <- 0.01
suba$N_Perv <- 0.1
suba$S_Imperv <- 0.05
suba$S_Perv <- 0.05
suba$PctZero <- 25
suba$RouteTo <- "OUTLET"

#used model defaults, N is for Manning's N, and S is for Depth of Depression Storage.
  
```

[INFILTRATION]					
;;Subcatchment  	MaxRate   	MinRate   	Decay     	DryTime   	MaxInfil  
;;--------------	----------	----------	----------	
1	3	0.5	4	7	0
```{r}
infil<-suba %>% select(Subcatchment)
infil$CurveNum <- 80
infil$Blank <- 0.5
infil$DryTime <- 7

#put in model defaults for now

```

[Polygons]		
;;Subcatchment  	X-Coord           	Y-Coord           
;;--------------	------------------	------------------
1	-19.96	8113.772
1	2594.81	7475.05
1	1077.844	5389.222
1	-1007.984	6926.148
1	-19.96	8133.733
```{r}
subc_xy <- subcatch%>% select(Name, X_Coord, Y_Coord) %>% rename(Subcatchment = Name) %>% distinct()
```

write csv file
```{r}
write.csv(subc ,"Wailupe_subc_SWMM.csv", row.names = FALSE)
write.csv(suba ,"Wailupe_suba_SWMM.csv", row.names = FALSE)
write.csv(subc_xy ,"Wailupe_poly_SWMM.csv", row.names = FALSE)
write.csv(infil ,"Wailupe_infil_SWMM.csv", row.names = FALSE)
```

