---
title: "results_maps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is code to create maps from the output files generated in Results_Graphs
*Run Results_Graphs before this!

# **Notes: 
-If you try to run the code all at once, the code may stop because of tmap_save() 
-tmap_save() is a function that allows you to export tmap outputs as images (such as .png)
-Something about this function stops the code, so you may need to run each code chunk manually 
-Even if the code returns an error, tmap_save() will still have generated an output file




Load packages
```{r, warning = FALSE, message = FALSE}
library(tidyverse) #For data wrangling
library(sf) #For shapefiles
library(tmap) #For mapmaking
library(tmaptools) #For mapmaking
library(here) #For loading shapefiles
library(janitor) #For cleaning names
```

Maps for dry storm hotspots
```{r, message=FALSE, warning = FALSE}
#Combine subcatchments outline with dry storm results
subcatch_dry <- read_sf(dsn = here("5.Results_Maps","shapefiles"), layer = "subcatch_outline") %>%
  st_transform(crs = 4326) %>% #Set coordinate system
  clean_names() %>% #Clean the names of columns
  select(subcatchment = objectid_1) %>% 
  merge(results_dry) %>% #Merge the subcatchment outlines to the SWMM results for the dry storm (from Results_Graphs)
  filter(subcatchment != "5") #Remove subcatchment 5 which is a mistake (overlaps another subcatchment)


#  Map total volume hotspots  
hotspots_dry_total <- tm_basemap("OpenStreetMap.Mapnik") +
  tm_shape(subcatch_dry, unit = "Miles") +
  tm_polygons("runoff_coeff", alpha = 0.8, palette = "Blues", style = "cont", n=8, title = "Runoff Coefficient") +
  tm_layout(title = "December 2010 storm", inner.margins=c(.05, .05, 0.1, .53), legend.position =  c(.6,.63), legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.6,.58), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.58,.51))
tmap_save(hotspots_dry_total, here("5.Results_Maps", "output_maps", "hotspots_dry_total.png"))

#  Map peak flow hotspots
hotspots_dry_peak <- tm_basemap("OpenStreetMap.Mapnik") +
  tm_shape(subcatch_dry, unit = "Miles") +
  tm_polygons("peak_runoff_cfs", alpha = 0.75, palette = "Greens", style = "cont", n=8, legend.hist = TRUE, title = "Peak Discharge (cfs)") +
  tm_layout(title = "December 2010 storm", inner.margins=c(.05, .05, 0.1, .53), legend.position =  c(.6,.35), legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.6,.61), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.58,.54))
tmap_save(hotspots_dry_peak, here("5.Results_Maps", "output_maps","hotspots_dry_peak.png"))

```

Maps for wet storm hotspots
```{r}
#Combine subcatchments outline with wet storm results
subcatch_wet <- read_sf(dsn = here("5.Results_Maps","shapefiles"), layer = "subcatch_outline") %>%
  st_transform(crs = 4326) %>% 
  clean_names() %>% 
  select(subcatchment = objectid_1) %>% 
  merge(results_wet) %>% 
   filter(subcatchment != "5")

#  Total volume hotspots  
hotspots_wet_total <- tm_basemap("OpenStreetMap.Mapnik") +
  tm_shape(subcatch_wet, unit = "Miles") +
  tm_polygons("runoff_coeff", alpha = 0.8, palette = "Blues", style = "cont", n=8, legend.hist = TRUE, title = "Runoff Coefficient") +
  tm_layout(title = "March 2009 storm", inner.margins=c(.05, .05, 0.1, .53), legend.position =  c(.6,.32), legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.6,.59), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.58,.52))
tmap_save(hotspots_wet_total, here("5.Results_Maps", "output_maps","hotspots_wet_total.png"))


#  Peak flow hotspots
hotspots_wet_peak <- tm_basemap("OpenStreetMap.Mapnik") +
  tm_shape(subcatch_wet, unit = "Miles") +
  tm_polygons("peak_runoff_cfs", alpha = 0.75, palette = "Greens", style = "cont", n=8, legend.hist = TRUE, title = "Peak Discharge (cfs)") +
  tm_layout(title = "March 2009 storm", inner.margins=c(.05, .05, 0.1, .53), legend.position =  c(.6,.27), legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.6,.54), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.58,.47))
tmap_save(hotspots_wet_peak, here("5.Results_Maps", "output_maps","hotspots_wet_peak.png"))
```

Maps for top 20 hotspots commonly found between the dry storm and the wet storm
```{r, warning = FALSE, message = FALSE}
#Top 20 of total volume
top20_dry_total <- subcatch_dry %>%  #Find top 20 hotspots for the dry storm
  arrange(-runoff_coeff) %>% 
  head(20)

top20_wet_total <- subcatch_wet %>% #Find top 20 hotspots for the wet storm
  arrange(-runoff_coeff) %>% 
  head(20)

#Find the top 20 hotspots that occur in both storms
common_total_vector <- as.data.frame(intersect(top20_dry_total$subcatchment, top20_wet_total$subcatchment)) 
colnames(common_total_vector) <- c("subcatchment")

#Select the top 20 hotspots 
common_total <- subcatch_dry %>% 
  mutate(hotspot = case_when(
  subcatchment == "21" | 
  subcatchment == "22" |
  subcatchment == "23" |
  subcatchment == "29" | 
  subcatchment == "40" | 
  subcatchment == "45" |
  subcatchment == "46" |
  subcatchment == "47" | 
  subcatchment == "49" | 
  subcatchment == "51" |
  subcatchment == "54" | 
  subcatchment == "59" | 
  subcatchment == "60" | 
  subcatchment == "63" | 
  subcatchment == "65" | 
  subcatchment == "67" | 
  subcatchment == "68" | 
  subcatchment == "71" | 
  subcatchment == "89" ~ "Hotspot"))

common_total$hotspot <- as.factor(common_total$hotspot)

#Create a map of the top 20 hotspots
top20_total_map <- tm_basemap("Hydda.Base") +
  tm_shape(common_total, unit = "Miles") +
  tm_polygons("hotspot", title = "Legend", textNA = "Subcatchment", palette="#045a8d", alpha = 0.8)+
  tm_layout(inner.margins=c(.05, .05, .05, .52), legend.position =  c(.56,.8),legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.56,.75), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.54,.68))

tmap_save(top20_total_map, here("5.Results_Maps", "output_maps","top20_total_map.png"))


#Top 20 of peak volume
top20_dry_peak <- subcatch_dry %>% 
  arrange(-peak_runoff_cfs) %>% 
  head(20)

top20_wet_peak <- subcatch_wet %>% 
  arrange(-peak_runoff_cfs) %>% 
  head(20)

common_peak_vector <- as.data.frame(intersect(top20_dry_peak$subcatchment, top20_wet_peak$subcatchment)) 
colnames(common_peak_vector) <- c("subcatchment")

common_peak <- subcatch_dry %>% 
  mutate(hotspot = case_when(
  subcatchment == "1" | 
  subcatchment == "2" |
  subcatchment == "3" |
  subcatchment == "7" | 
  subcatchment == "11" | 
  subcatchment == "23" |
  subcatchment == "28" |
  subcatchment == "38" | 
  subcatchment == "42" | 
  subcatchment == "58" |
  subcatchment == "62" | 
  subcatchment == "63" | 
  subcatchment == "74" | 
  subcatchment == "75" | 
  subcatchment == "76" | 
  subcatchment == "78" | 
  subcatchment == "79" | 
  subcatchment == "89" | 
  subcatchment == "94" ~ "Hotspot"))

common_peak$hotspot <- as.factor(common_peak$hotspot)

top20_peak_map <- tm_basemap("Hydda.Base") +
  tm_shape(common_peak, unit = "Miles") +
  tm_polygons("hotspot", title = "Legend", textNA = "Subcatchment", palette="#006d2c", alpha = 0.8)+
  tm_layout(inner.margins=c(.05, .05, .05, .52), legend.position =  c(.56,.8),legend.title.size = 1.4, legend.text.size = 1) +
  tm_text("subcatchment", size = 0.3) +
  tm_scale_bar(position = c(.56,.75), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  tm_compass(position = c(.54,.68))

tmap_save(top20_peak_map, here("5.Results_Maps", "output_maps","top20_peak_map.png"))

```


Export the results as shapefiles (e.g. For use in other map software like ArcGIS or Google Earth Pro)
```{r}
#Keep the common hotspots only
common_peak_only <- subcatch_dry %>% 
  filter(
  subcatchment == "1" | 
  subcatchment == "2" |
  subcatchment == "3" |
  subcatchment == "7" | 
  subcatchment == "11" | 
  subcatchment == "23" |
  subcatchment == "28" |
  subcatchment == "38" | 
  subcatchment == "42" | 
  subcatchment == "58" |
  subcatchment == "62" | 
  subcatchment == "63" | 
  subcatchment == "74" | 
  subcatchment == "75" | 
  subcatchment == "76" | 
  subcatchment == "78" | 
  subcatchment == "79" | 
  subcatchment == "89" | 
  subcatchment == "94" )

#Export as .shp
st_write(common_peak_only, here("5.Results_Maps", "output_shapefiles", "common_peak.shp"))


#Keep the common hotspots only
common_total_only <- subcatch_dry %>% 
  filter(
  subcatchment == "21" | 
  subcatchment == "22" |
  subcatchment == "23" |
  subcatchment == "29" | 
  subcatchment == "40" | 
  subcatchment == "45" |
  subcatchment == "46" |
  subcatchment == "47" | 
  subcatchment == "49" | 
  subcatchment == "51" |
  subcatchment == "54" | 
  subcatchment == "59" | 
  subcatchment == "60" | 
  subcatchment == "63" | 
  subcatchment == "65" | 
  subcatchment == "67" | 
  subcatchment == "68" | 
  subcatchment == "71" | 
  subcatchment == "89")

#Export as .shp
st_write(common_total_only,  here("5.Results_Maps", "output_shapefiles","common_total.shp"))
```
