---
title: "Soil Curve Numbers"
author: "Eleonore Durand"
date: "May 13, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
#load necessary packages

library(tidyverse)

```


```{r, message = FALSE, fig.show='hide'}
soil_union <- read_csv("soil_land_union_maunalua.csv") %>% 
  select(OBJECTID_1, hydgrpdcd, Class_Name, Shape_Length, Shape_Area) %>% 
  filter(hydgrpdcd != "NA")

unique(soil_union$Class_Name) # get all the class names 


```

See methodology for further explanation as to how we found the curve number for each land class, but here is the table for ease of coding: 

Land Cover	               Hydrologic group A	Hydrologic group B	Hydrologic group C	Hydrologic group D
Grassland: poor condition	          68	              79	              86	                89
Unconsolidated shore	              0	                0	                0	                   0
Bare Land/Bare soil	                77                86	              91                	94
Open Space Developed- good	        39              	61              	74                	80
Evergreen forest- fair	            36	              60              	73                	79
Scrub Shrub	                        36	              42	              55	                62
Open Water	                        0               	0               	0	                   0
Impervious surface- like shrubland	36	              42	              55	                62
Palustrine Scrub Shrub wetland
(woody wetland )	                  86	              86	              86	                86
Palustrine Forested wetland 
(woody wetland )	                  86	              86	              86	                86
Palustrine Aquatic Bed	            NA	              NA	              NA	                NA
Estuarine Emergent wetland	        80	              80	              80	                80
Palustrine emergent wetland	        80	              80	              80	                80
Pasture/Hay	                        40	              61	              73	                79
Unclassified - here: open water	    0	                0	                0	                  0
Cultivated Land	                    62	              74	              82	                86


```{r}
#lets add the curve numbers 

curve_numbers_maunalua <- soil_union %>% 
  mutate(CN= 
           case_when(Class_Name =="Grassland" & hydgrpdcd== "A" ~ 68, 
                     Class_Name =="Grassland" & hydgrpdcd== "B" ~ 79, 
                     Class_Name =="Grassland" & hydgrpdcd== "C" ~ 86, 
                     Class_Name =="Grassland" & hydgrpdcd== "D" ~ 89, 
                     Class_Name =="Unconsolidated Shore"~ 0, 
                     Class_Name =="Bare Land" & hydgrpdcd== "A" ~ 77, 
                     Class_Name =="Bare Land" & hydgrpdcd== "B" ~ 86, 
                     Class_Name =="Bare Land" & hydgrpdcd== "C" ~ 91, 
                     Class_Name =="Bare Land" & hydgrpdcd== "D" ~ 94, 
                     Class_Name =="Open Space Developed" & hydgrpdcd== "A" ~ 39, 
                     Class_Name =="Open Space Developed" & hydgrpdcd== "B" ~ 61,
                     Class_Name =="Open Space Developed" & hydgrpdcd== "C" ~ 74,
                     Class_Name =="Open Space Developed" & hydgrpdcd== "D" ~ 80, 
                     Class_Name =="Evergreen" & hydgrpdcd== "A" ~ 36,
                     Class_Name =="Evergreen" & hydgrpdcd== "B" ~ 60,
                     Class_Name =="Evergreen" & hydgrpdcd== "C" ~ 73,
                     Class_Name =="Evergreen" & hydgrpdcd== "D" ~ 79,
                     Class_Name =="Scrub Shrub" & hydgrpdcd== "A" ~ 39,
                     Class_Name =="Scrub Shrub" & hydgrpdcd== "B" ~ 42,
                     Class_Name =="Scrub Shrub" & hydgrpdcd== "C" ~ 55,
                     Class_Name =="Scrub Shrub" & hydgrpdcd== "D" ~ 62,
                     Class_Name =="Open Water" ~ 0,
                     Class_Name =="Impervious Surface" & hydgrpdcd== "A" ~ 39,
                     Class_Name =="Impervious Surface" & hydgrpdcd== "B" ~ 42,
                     Class_Name =="Impervious Surface" & hydgrpdcd== "C" ~ 55,
                     Class_Name =="Impervious Surface" & hydgrpdcd== "D" ~ 62,
                     Class_Name =="Estuarine Emergent Wetland" ~ 80, 
                     Class_Name =="Estuarine Scrub Shrub Wetland" ~ 86,
                     Class_Name =="Estuarine Forested Wetland" ~ 86,
                     Class_Name =="Cultivated Land" & hydgrpdcd== "A" ~ 62,
                     Class_Name =="Cultivated Land" & hydgrpdcd== "B" ~ 74,
                     Class_Name =="Cultivated Land" & hydgrpdcd== "C" ~ 82,
                     Class_Name =="Cultivated Land" & hydgrpdcd== "D" ~ 86
             
           ))

```

Then export this file to a csv and merge back the CN column to your attribute file for your union land uses/soil type. Double check that you are joining by the right column, here OBJECTID_1. 
